---
title: "Data"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---


The dataset used in this project was sourced from Kaggle, where it was provided as a CSV file. You can explore or download the dataset directly from Kaggle by visiting [this link](https://www.kaggle.com/code/kinguistics/exploring-meteorite-landing-sites/input).


# Clean data

As part of the data cleaning, we standardized the column names by converting them all to lowercase for consistency and ease of use in future analysis. We then filtered the dataset to include only records from the years 1970 to 2013, ensuring that the analysis focused on a more relevant and manageable timeframe. Next, we identified and removed the "fall" column since all its entries contained the same value. Rows with missing latitude or longitude values were also excluded to maintain the accuracy and reliability of geographic data. Additionally, we enriched the dataset by creating a new "continent" column. This was achieved by classifying each record based on its `geo_location` attribute, which allows for more detailed geographic insights in the analysis. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(janitor)
library(ggplot2)
library(maps)
library(plotly)
library(rnaturalearth)
library(sf)
library(RColorBrewer)
library(car)
```

```{r}
meteorite_data <- read.csv("data/meteorite-landings.csv")

meteorite_data <- meteorite_data %>% 
  clean_names() %>%
  filter(year >= 1970 & year <= 2013) %>% 
  select(-fall)

meteorite_data_cleaned <- meteorite_data %>%
  filter(!is.na(reclat) & !is.na(reclong)) %>%
  filter(!is.na(mass) & !is.na(year)) %>%
  filter(mass > 0)
```


# Glimpse of data

```{r}
meteorite_sf <- st_as_sf(meteorite_data_cleaned, coords = c("reclong", "reclat"), crs = 4326, remove = FALSE)

continents <- ne_countries(scale = "medium", returnclass = "sf") %>%
  group_by(continent) %>%
  summarise()

meteorite_data_with_continents <- st_join(meteorite_sf, continents["continent"], join = st_intersects)

meteorite_data_with_continents <- meteorite_data_with_continents %>%
  mutate(continent = replace_na(continent, "Ocean"))

head(meteorite_data_with_continents)
```

The dataset is a simple feature collection comprising 6 features and 10 fields. The fields include key attributes such as the name, id, mass, year, latitude, and longitude of meteorite landings, along with a derived continent column.

# Statistics of Variables

```{r}
summary(meteorite_data_cleaned)
```

The dataset contains 35,882 entries. The mass variable ranges from 0 to 4,000,000 grams, with a mean of 1,484 grams and a median of 25 grams, indicating significant skewness. The year variable spans from 1970 to 2013, with a median year of 1997. Latitude and longitude values reflect global coverage, while geographic and meteorite classification details enrich the dataset.

# Distribution of Mass

```{r}
meteorite_data_cleaned %>%
  ggplot(aes(x = mass)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") +
  scale_x_log10() +
  labs(
    title = "Distribution of Meteorite Mass",
    x = "Mass (log scale)",
    y = "Count"
  )
```

The mass distribution is heavily skewed due to the presence of extremely large meteorites. To normalize this skew, the log transformation of the mass was applied, resulting in a distribution resembling a normal distribution. This transformation facilitates better visualization and analysis.

# Numbers of landings through Year

```{r}
meteorite_data_cleaned %>%
  ggplot(aes(x = year)) +
  geom_bar(fill = "coral", color = "black") +
  labs(
    title = "Meteorite Landings Over Time",
    x = "Year",
    y = "Number of Landings"
  )
```

Meteorite landings were distributed unevenly across the years, with a noticeable increase in records toward the later part of the timeframe. This could be attributed to improved detection and recording technologies. A time-series plot provides insights into the frequency trends over the years.

# Final Dataset Description

`name`: The name of the meteorite landing.

`id`: A unique identifier for each meteorite.

`nametype`: The classification of the name type (e.g., "Valid").

`recclass`: The classification of the meteorite based on its composition.

`mass`: The mass of the meteorite in grams.

`year`: The year the meteorite landed or was discovered.

`reclat`: The latitude coordinate of the meteorite landing site.

`reclong`: The longitude coordinate of the meteorite landing site.

`geo_location`: A formatted string combining the latitude and longitude of the meteorite landing.

`continent`: The continent where the meteorite landed, derived from the latitude and longitude.